---
title: "Pleiotropy for quantitive traits: pleio"
author: "Dan Schaid, Jason Sinnwell"
date: "`r format(Sys.time(),'%d %B, %Y')`"
output: 
  rmarkdown:::pdf_document:
    toc: false
    toc_depth: 3
---

Example Data
=============

The pleio package contains a pre-made simulated dataset with multiple quantitative traits 
simulated from a multivariate normal distribution with common correlation
structure, correlation of 0.5, and genotypes simulated based on minor allele
frequency of 0.2, and assumes that the regression coefficients (beta's) for traits 2 and 3 have
non-zero values, while all other traits are not asssociated with dose
of minor allele.

Here, we load the simulated dataset and show
```{r, loaddat}

## load package and dataset
require(pleio)
data(pleio.qsim)

## preview simulated data
head(y)
table(geno)
```

## Sequential Pleiotropy Tests

The *pleio.q.sequential* function is a high-level way to perform sequential tests of the number of traits (and which traits) are 
associated with a genotype. The algorithm starts with testing the usual multivariate null hypthothesis that all beta's are zero. If this test rejects, because the p-value is less than a user-spiecifed threshold, then allow one beta to be non-zero in order to test whether the remaining beta's $= 0$. If the test of one non-zero beta rejects, then allow two non-zero beta's, and continue this sequential testing until the p-value for a test is greater than the specified threshold. When the sequential testing stops, one can conclude that the final model contains the non-zero beta's, while all other beta's are inferred to be zero.  Below we run two functions, *pleio.q.fit*, which performs pre-calculations on the models to be tested, and *pleio.q.sequential*, which performs the sequential pleiotropy tests.

The final result lists the indices of the non-zero beta's (the indices of the traits associated with a genotype), and the p-value that tests the fit of the final model. A p-value greater than the threshold is expected for the final model, showing that the final model fits the data well. For this example, the
sequential approach stopped at 2 traits because the p-value is greater than the
*pval.threshold* argument given of 0.05.

```{r, sequentialfit}
fit <- pleio.q.fit(y, geno)

test.seq <- pleio.q.sequential(fit, pval.threshold=.05)
test.seq
```

## Equivalent Steps to Sequential Fit

The sequential steps above can be performed with more user control using *pleio.q.test*, with
*count.nonzero.beta* as the number of non-zero betas for the null hypothesis. The result of
*pleio.q.test* contains the global test statistic, degrees of freedom (df), p-value for testing the
model, the index of the non-zero beta's in the model, and a data.frame called "tests" that contains the tests performed for the null hypothesis models (i.e., the indices of the non-zero beta's and the corresponding statistic, tk, for each model). For $m$ traits, and $k = count.nonzero.beta$, there are m-choose-k models in the set that are considered in the null hypothesis,
and the minimum tk test statistic over the set provides the global test statistic reported.

```{r, fitsteps}
test0 <- pleio.q.test(fit, count.nonzero.beta = 0)
test0

test1 <- pleio.q.test(fit, count.nonzero.beta = 1)
test1

test2 <- pleio.q.test(fit, count.nonzero.beta = 2)
test2

```

